---
title: '251107'
author: "TF"
date: "2025-11-07"
output: html_document
---

# Setup

## Directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

dir_wk <- "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript"
knitr::opts_knit$set(root.dir = dir_wk)
```

## Import modules

```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
```

## Important files

All cells (sham + vsg_new&old) + Azi annotation + joined layers + UMAP\
(all cell type determined by Azimuth annotation)

```         
# Integrated + modified cell types
sobj_all_2, file = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/251029_anno_integrated_projVSG.rds"
```

Integrated endocrine cells (sham+vsg) + *Gast*-positive cells (sham) + UMAP clusters + endo-subtype annotation (markers, manual)

```         
sobj_0, file = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/251024_anno_umap_endo+gast.rds"

# Integrated, annotated based on marker expression & seurat_clusters
sobj_1, file = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/251030_anno_integrated_endocrine.rds"
```

## Read files

```{r}

file_path <- "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/251029_anno_integrated_projVSG.rds"

sobj_all_2 <- readRDS(file = file_path)
```

## Save files

```{r}
saveRDS(sobj, file = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/251030_anno_integrated_endocrine.rds")
```

## Change Font Color

```         
<span style="color: #0066CC;">ooo</span>
[ooo]{style="color: #0066CC;"}
```

# Visualize all cell types

(sobj_all_2) - "concise_Main_cell_type" "concise_Main_cell_ID"

## [Cell types]{style="color: #0066CC;"}

### Colors

```         
#E41A1C
#4E1DFA
#00AD59
#FEC001
#BA0503
#FE7E02
#B11381
#0B2A67
#00712D
#D63156
#00BFC4
#7B3294
#FB8072
#008080
#CF76FA
```

```{r}
sobj <- subset(
  sobj_all_2,
  subset = concise_Main_cell_type != "Non-stomach"
)

table(sobj$concise_Main_cell_type)
table(sobj$concise_Main_cell_ID)
```

Colors assigned by concise_Main_cell_type

### 1. Legends: Cell types

```{r}
my_colors <- c(
  "Chief cells" = "#E41A1C",
  "Endocrine cells" = "#4E1DFA",
  "ENS cells" = "#00AD59",
  "Fibroblasts" = "#FEC001",
  "Gastric epithelial cells" = "#BA0503",
  "Gastric mucous cells" = "#FE7E02",
  "Lymphatic endothelial cells" = "#B11381",
  "Lymphoid cells" = "#0B2A67",
  "Mesothelial cells" = "#00712D",
  "Mural cells" = "#D63156",
  "Myeloid cells" = "#00BFC4",
  "Parietal cells" = "#7B3294",
  "Tuft cells" = "#FB8072",
  "Vascular endothelial cells" = "#008080"
  # "Non-stomach" = "snow2"
)

#  create a number index for cell types
cell_type_order <- names(my_colors)
#  match() 会返回 sobj$concise_cell_type 中每个值
sobj$concise_Main_cell_ID <- match(sobj$concise_Main_cell_type, cell_type_order)

# UMAP clusters - joined
p1 <- DimPlot(sobj, 
        group.by = "concise_Main_cell_type"
        ) +
  ggtitle("All Cell Types (Joined)") +
  scale_color_manual(values = my_colors)
p1

ggsave(
  filename = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/rplot_251107_allcelltype_joined_ShamVSG.jpeg",
  plot = p1,
  width = 10,
  height = 8,
  units = "in",
  dpi = 300
)

# UMAP clusters - split by treatment (Sham/VSG)
p2 <- DimPlot(sobj, 
        group.by = "concise_Main_cell_type",
        split.by = "treatment"
        ) +
  ggtitle("All Cell Types (Split)") +
  NoLegend() +
  scale_color_manual(values = my_colors)
p2

ggsave(
  filename = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/rplot_251107_allcelltype_split_ShamVSG.jpeg",
  plot = p2,
  width = 16,
  height = 8,
  units = "in",
  dpi = 300
)
```

### 2. Legends: number indices

```{r}
my_colors <- c(
  "1" = "#E41A1C",
  "2" = "#4E1DFA",
  "3" = "#00AD59",
  "4" = "#FEC001",
  "5" = "#BA0503",
  "6" = "#FE7E02",
  "7" = "#B11381",
  "8" = "#0B2A67",
  "9" = "#00712D",
  "10" = "#D63156",
  "11" = "#00BFC4",
  "12" = "#7B3294",
  "13" = "#FB8072",
  "14" = "#008080"
  # "15" = "snow2"
)

# UMAP clusters - joined
p3 <- DimPlot(sobj, 
        group.by = "concise_Main_cell_ID",
        label = TRUE
        ) +
  ggtitle("All Cell Types (Joined)") +
  scale_color_manual(values = my_colors)
p3

ggsave(
  filename = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/rplot_251107_allcelltype_joined_ShamVSG_legendnumber.jpeg",
  plot = p3,
  width = 8,
  height = 8,
  units = "in",
  dpi = 300
)

# UMAP clusters - split by treatment (Sham/VSG)
p4 <- DimPlot(sobj, 
        group.by = "concise_Main_cell_ID",
        split.by = "treatment",
        label = TRUE
        ) +
  ggtitle("All Cell Types (Split)") +
  scale_color_manual(values = my_colors)
p4

ggsave(
  filename = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/rplot_251107_allcelltype_split_ShamVSG_legendnumber.jpeg",
  plot = p4,
  width = 16,
  height = 8,
  units = "in",
  dpi = 300
)
```

## [Marker-expressing cells highlighted]{style="color: #0066CC;"}

### Single target gene

```         
Npc2
Pcyox1
Apoa4
Manf
Pepc    Pck1 is the correct gene name
Cblif
Pnliprp2
Krt10
```

```{r}
library(patchwork)
marker <- "Krt10"
threshold <- 0
aspect_ratio <- 1.0

plot_list <- FeaturePlot(sobj, 
                  features = marker,
                  split.by = "treatment",
                  combine = FALSE,
                  min.cutoff = threshold)
# Stack them manually: [1] on top of [2]
p2_vertical <- plot_list[[1]] / plot_list[[2]]

p2_vertical <- p2_vertical & theme(aspect.ratio = aspect_ratio)

p2_final <- p2_vertical & 
  theme(
    # 1. Moves "Sham" and "VSG" from the side to the top
    strip.position = "top", 
    
    # 2. Removes the grey/black box *behind* the text
    strip.background = element_blank(),
    
    # 3. (Optional) Make the text bold and larger
    strip.text = element_text(face = "bold", size = 12) 
  )

p2_final
```

```{r}
output_dir <- "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/"

dynamic_filename <- paste0(
    output_dir, 
    "rplot_251120_allcell_marker_highlighted_",
    marker,
    "_",
    threshold,
    ".jpeg"
  )
 
ggsave(
  filename = dynamic_filename,
  plot = p2_final,
  width = 10,
  height = 8,
  units = "in",
  dpi = 300
)
```

### Gene list

```         
CHIA    Chia1
ATP4A   Atp4a
ATP4B   Atp4b
PDIA2   Pdia2
CLIC6   Clic6
NUCB2   Nucb2
FURIN   Furin
SAR1B   Sar1b
GHRL    Ghrl
SND1    Snd1
ERO1A   Ero1a
HDGF    Hdgf
```

```         
Pla2g1b
Clps
Pgc
```

```         
Npc2
Pcyox1
Apoa4
Manf
Pepc    Pck1 ?
Cblif
Pnliprp2
Krt10
```

### For loop

```{r}
library(patchwork)

# --- 1. gene_list ---
gene_list <- c(
  "Chia1", 
  "Atp4a", 
  "Atp4b", 
  "Pdia2", 
  "Clic6", 
  "Nucb2", 
  "Furin", 
  "Sar1b",
  "Ghrl",
  "Snd1",
  "Ero1a",
  "Hdgf"
  ) 

# --- 2. output dir ---
output_dir <- "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/"

# --- 3. for loop 'gene_list' ---
for (current_gene in gene_list) {
  
  # does current_gene exist?
  if (!current_gene %in% rownames(sobj)) {
    
    print(paste("---!!! Skipping:", current_gene, "(Gene not found in sobj) ---"))
    next
  }
  
  
  print(paste("--- Generating plot for:", current_gene, "---"))
  
  plot_list <- FeaturePlot(sobj, 
                           features = current_gene, 
                           split.by = "treatment",
                           combine = FALSE)
  
  # Stack them manually: [1] on top of [2]
  p2_vertical <- plot_list[[1]] / plot_list[[2]]
  
  p2_vertical <- p2_vertical & theme(aspect.ratio = 1.5)
  
  p2_final <- p2_vertical & 
    theme(
      strip.position = "top", 
      strip.background = element_blank(),
      strip.text = element_text(face = "bold", size = 12) 
    )
  
  # --- 4. dynamic filenames & save ---
  dynamic_filename <- paste0(
    output_dir, 
    "rplot_251107_allcell_marker_highlighted_",
    current_gene, # <-- 动态变量
    ".jpeg"
  )
  
  ggsave(
    filename = dynamic_filename, 
    plot = p2_final,
    width = 10,
    height = 8,
    units = "in",
    dpi = 300
  )
  
}

print("--- All plots saved. ---")
```

# Visualize endocrine cells

(sobj_0) - "endo_manual"

```{r}
sobj <- subset(
  sobj_0,
  subset = endo_manual != "None"
)

sobj$endo_manual[sobj$endo_manual == "Gast (G cell)"] <- "G cells"
sobj$endo_manual[sobj$endo_manual == "Ghrl (X/A cell)"] <- "X/A-like cells"
sobj$endo_manual[sobj$endo_manual == "Hdc (ECL cell)"] <- "ECL cells"
sobj$endo_manual[sobj$endo_manual == "Sst (D cell)"] <- "D cells"
sobj$endo_manual[sobj$endo_manual == "Tph1 (EC cell)"] <- "EC cells"

table(sobj$endo_manual)
```

### Color

```         
#BA0503
#FE7E02
#00AD59
#0B2A67
#4E1DFA
```

## [Cell types]{style="color: #0066CC;"}

```{r}

# my_colors_1 <- c(
#   "D cells" = "#BA0503",
#   "EC cells" = "#FE7E02",
#   "ECL cells" = "#00AD59",
#   "G cells" = "#0B2A67",
#   "X/A-like cells" = "#4E1DFA"
# )

my_colors_2 <- c(
  "D cells (Sst)" = "#FF7F29",
  "EC cells (Tph1)" = "#C7599B",
  "ECL cells (Hdc)" = "#94DE5E",
  "G cells (Gast)" = "#29A5EB",
  "X/A-like cells (Ghrl)" = "#FFC847"
)

# UMAP clusters - joined
p1 <- DimPlot(sobj, 
        group.by = "endo_manual"
        ) +
  ggtitle("Endocrine Cell Types (Joined)") +
  scale_color_manual(values = my_colors_2)
p1

ggsave(
  filename = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/rplot_251107_endocrine_joined_ShamVSG_clr2.jpeg",
  plot = p1,
  width = 10,
  height = 8,
  units = "in",
  dpi = 300
)

# UMAP clusters - split by treatment (Sham/VSG)
p2 <- DimPlot(sobj, 
        group.by = "endo_manual",
        split.by = "treatment"
        ) +
  ggtitle("Endocrine Cell Types (Split)") +
  NoLegend() +
  scale_color_manual(values = my_colors_2)
p2

ggsave(
  filename = "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/rplot_251107_endocrine_split_ShamVSG_clr2.jpeg",
  plot = p2,
  width = 16,
  height = 8,
  units = "in",
  dpi = 300
)
```

## [Stacked violin plots for marker gene expression]{style="color: #0066CC;"}

<https://divingintogeneticsandgenomics.com/post/stacked-violin-plot-for-visualizing-single-cell-data-in-seurat/>

```{r}
library(Seurat)
library(patchwork)
library(ggplot2)

## remove the x-axis text and tick
## plot.margin to adjust the white space between each plot.
## ... pass any arguments to VlnPlot in Seurat
modify_vlnplot <- function(obj, 
                          feature, 
                          pt.size = 0, 
                          group.by = NULL,  # <-- MODIFIED: Add group.by argument
                          plot.margin = unit(c(-0.75, 0, -0.75, 0), "cm"),
                          ...) {
  p <- VlnPlot(obj, 
              features = feature, 
              group.by = group.by, # <-- MODIFIED: Pass to VlnPlot
              pt.size = pt.size, ... )  + 
    xlab("") + ylab(feature) + ggtitle("") + 
    theme(legend.position = "none", 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(), 
          axis.title.y = element_text(size = rel(1), angle = 0), 
          axis.text.y = element_text(size = rel(1)), 
          plot.margin = plot.margin ) 
  return(p)
}

## extract the max value of the y axis
extract_max <- function(p){
  ymax <- max(ggplot_build(p)$layout$panel_scales_y[[1]]$range$range)
  return(ceiling(ymax))
}


## main function
StackedVlnPlot <- function(obj, features,
                          pt.size = 0, 
                          group.by = NULL,  # <-- MODIFIED: Add group.by argument
                          plot.margin = unit(c(-0.75, 0, -0.75, 0), "cm"),
                          ...) {
  
  plot_list <- purrr::map(features, function(x) {
    modify_vlnplot(obj = obj,
                   feature = x, 
                   group.by = group.by, # <-- MODIFIED: Pass to helper
                   ...)
    })
  
  # Add back x-axis title to bottom plot. patchwork is going to support this?
  plot_list[[length(plot_list)]]<- plot_list[[length(plot_list)]] +
    theme(axis.text.x=element_text(angle = 45), axis.ticks.x = element_line())
  
  # change the y-axis tick to only max value 
  ymaxs <- purrr::map_dbl(plot_list, extract_max)
  plot_list<- purrr::map2(plot_list, ymaxs, function(x,y) x + 
                           scale_y_continuous(breaks = c(0, y)) + 
                           expand_limits(y = y))

  p <- patchwork::wrap_plots(plotlist = plot_list, ncol = 1)
  
  p <- p & theme(aspect.ratio = 0.2)
  
  return(p)
}
```

### Plot using my data

```         
my_colors_2 <- c(
  "D cells" = "#FF7F29",
  "EC cells" = "#C7599B",
  "ECL cells" = "#94DE5E",
  "G cells" = "#29A5EB",
  "X/A-like cells" = "#FFC847"
)
```

```{r}
features<- c("Ghrl", "Gast", "Hdc", "Tph1", "Sst")

my_colors <- c(
  "X/A-like cells" = "#27e3c7",
  "G cells" = "#c1f195",
  "D cells" = "#55832f",
  "EC cells" = "#8ab961",
  "ECL cells" = "#354b46"
)

p3 <- StackedVlnPlot(obj = sobj, 
               features = features, 
               group.by = "endo_manual_1030",
               cols = my_colors)
p3
```

## [Marker-expressing cells highlighted]{style="color: #0066CC;"}

### Gene list

```         
"Ghrl", "Gast", "Hdc", "Tph1", "Sst"
```

### For Loop

```{r}
library(patchwork)

# --- 1. gene_list ---
gene_list <- c("Ghrl", "Gast", "Hdc", "Tph1", "Sst") 

# --- 2. output dir ---
output_dir <- "/cluster/home/tfeng02/R/202510_TF_VSG_analysis_VSGmanuscript/"

# --- 3. for loop 'gene_list' ---
for (current_gene in gene_list) {
  
  # does current_gene exist?
  if (!current_gene %in% rownames(sobj)) {
    
    print(paste("---!!! Skipping:", current_gene, "(Gene not found in sobj) ---"))
    next
  }
  
  
  print(paste("--- Generating plot for:", current_gene, "---"))
  
  plot_list <- FeaturePlot(sobj, 
                           features = current_gene, 
                           split.by = "treatment",
                           combine = FALSE)
  
  # Stack them manually: [1] on top of [2]
  p2_vertical <- plot_list[[1]] / plot_list[[2]]
  
  p2_vertical <- p2_vertical & theme(aspect.ratio = 1.5)
  
  p2_final <- p2_vertical & 
    theme(
      strip.position = "top", 
      strip.background = element_blank(),
      strip.text = element_text(face = "bold", size = 12) 
    )
  
  # --- 4. dynamic filenames & save ---
  dynamic_filename <- paste0(
    output_dir, 
    "rplot_251107_endocrine_marker_highlighted_",
    current_gene, # <-- 动态变量
    ".jpeg"
  )
  
  ggsave(
    filename = dynamic_filename, 
    plot = p2_final,
    width = 10,
    height = 8,
    units = "in",
    dpi = 300
  )
  
}

print("--- All plots saved. ---")
```
